---
title: "Prediction Interval"
author: "Ankit Tyagi"
date: "1/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Prediction Intervals :-
A prediction interval is an interval within which we expect $y_t$ to lie with a specified probability. Prediction intervals are required to understand how much uncertainity is associated with the point forecast.

If the forecast errors are normally distributed, a 95% prediction interval is given by :-
$$y_{T+h|T} \pm 1.96\ est(σ_{h})$$
where h is forecast horizon.

1. One step Prediction usually have forecast error close to the residuals. In case the method do not have a paramter to estimate they are same.
2. Multi step prediction interval increase as the forecast horizon increase.

The estimate of $σ_{h}$ for different methods are :-
Mean Forecasts :- $σ \sqrt{1+1/T}$.
Naive Forecasts :- $σ \sqrt{h}$.
Seasonal Naive Forecasts :- $σ \sqrt{k +1}$.
Drift Forecasts :- $σ \sqrt{h(1+h/T)}$.

Prediction intervals for naive methods on Google Price index data
```{r}
library(fpp2)
naive(goog200)
```

It is clear that as we are moving up in the forecast horizon the prediction intervals are getting wider. This can also be seen in the graph below.
```{r}
autoplot(naive(goog200))
```

Bootstrap Prediction interval :- If the assumption of normal distribution of forecast error is  unreasonable, then we can use bootstrapping that only assumes that the forecast errors are uncorrelated.

Forecast error 
$$e_t = y_t - est(y_{t|t-1})$$
$$=> y_t = est(y_{t|t-1}) + e_t$$
We can simulate the next observation of a time series using 
$$y_{T+1} = est(y_{T+1|T}) + e_{T+1}$$
where $e_{T+1}$ is the unknown future error.
Assuming future errors are uncorrelated from the past error, we can replace $e_{T+1}$  by sampling from the collection of error we have seen in the past(i.e., residuals). We can repeat this process to obtain 
$$y_{T+2} = est(y_{T+2|T+1}) + e_{T+2}$$

where $e_{T+1}$ is another draw from the collection of residuals.
Doing this repeatedly we obtain many possible futures. Then we can compute prediction intervals by calculating percentiles for each forecast horizon. The result is called a bootstrapped prediction interval.

```{r}
naive(goog200, bootstrap = TRUE)
```

If the normal distribution assumption holds then bootstrap PI and PI based on normal distribution are similar.

If a transformation has been used, then the prediction interval should be computed on the transformed scale, and the end points back-transformed to give a prediction interval on the original scale. This approach preserves the probability coverage of the prediction interval, although it will no longer be symmetric around the point forecast.
