---
title: "Prediction Interval"
author: "Ankit Tyagi"
date: "1/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Prediction Intervals :-
A prediction interval is an interval within which we expect $y_t$ to lie with a specified probability. Prediction intervals are required to understand how much uncertainity is associated with the point forecast.

If the forecast errors are normally distributed, a 95% prediction interval is given by :-
$$y_{T+h|T} \pm 1.96\ est(σ_{h})$$
where h is forecast horizon.

1. One step Prediction usually have forecast error close to the residuals. In case the method do not have a paramter to estimate they are same.
2. Multi step prediction interval increase as the forecast horizon increase.

The estimate of $σ_{h}$ for different methods are :-
Mean Forecasts :- $σ \sqrt{1+1/T}$.
Naive Forecasts :- $σ \sqrt{h}$.
Seasonal Naive Forecasts :- $σ \sqrt{k +1}$.
Drift Forecasts :- $σ \sqrt{h(1+h/T)}$.

Prediction intervals for naive methods on Google Price index data
```{r}
library(fpp2)
naive(goog200)
```

It is clear that as we are moving up in the forecast horizon the prediction intervals are getting wider. This can also be seen in the graph below.
```{r}
autoplot(naive(goog200))
```

