---
title: "5(b). Trend and Seasonal Methods"
author: "Ankit"
date: "2/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Holt's linear trend method :-
This method extends the simple exponential smoothing to allow the forecasting of data with a trend.

This method has 2 smoothing equations and 1 forecast equations:-

**Forecast Equation :- ** $\hat y_{t+h|t} = l_t + hb_t$.

**Level Equation :- ** $l_t = αy_t + (1-α)(l_{t-1} + b_{t-1})$.

**Trend Equation :- ** $b_t = β^* (l_t-t_{t-1}) + (1-β^*)b_{t-1}$.

$l_t$ = estimate of level at time t.
$b_t$ = estimate of trend at time t. 

α is smooting parameter for level and $β^*$ is the smoothing parameter for trend. Both takes values between 0 and 1.

As with simple exponential smoothing, the level $l_t$ is the weighted average of the observation $y_t$ and the one-step training forecast of for time t, here given by $l_{t-1} + b_{t-1}$.
The trend component is the weighted average of estimated trend at time t based on $l_t - l_{t-1}$ and previous estimate of trend $b_{t-1}$.

**Example :- Air Passanger data**
```{r}
library(fpp2)
autoplot(window(ausair, start = 1990)) +
  xlab("Year") + ylab("") 
```

The timeplot shows that there is an increasing trend in the data.
The smoothing parameters α, $β^*$, $l_o$, $b_o$ are estimated by minimising the SSE for the one-step training errors.
We are fitting a holt's method on this data.
```{r}
air = window(ausair, start = 1990)
fc = holt(air, h = 5)
fc$model
```

The value of α = 0.8302.
$β^*$ = 0.0001.
The very small value of $β^*$ means that slope hardly changes over time.

The Holt's method assumes that the trend is constantly increasing or decreasing with time. This might leads to over forecast for a longer forecast horizons.

###Damped Trend Methods :-
This method introduces a parameter(ϕ) that "dampens" the trend to a flat line sometime in the future.

**Forecast Equation :- ** $\hat y_{t+h|t} = l_t + (ϕ + ϕ^2 + ϕ^3 + ...... ϕ^h)b_t$.

**Level Equation :- ** $l_t = αy_t + (1-α)(l_{t-1} + ϕb_{t-1})$.

**Trend Equation :- ** $b_t = β^* (l_t-t_{t-1}) + (1-β^*)ϕb_{t-1}$.

If ϕ = 1, the method is equal to Holt's linear method. For values between 0 and 1, ϕ dampens the trend so that it approaches a constant some time in the future. In fact, the forecasts converge to $l_T + ϕb_T/(1-ϕ)$ as h -> ∞.for any value 0<ϕ<1.This means that short-run forecasts are trended while long-run forecasts are constant.

```{r}
fc = holt(air,h =15)
fc2 = holt(air, damped = TRUE, phi = 0.9, h = 15)
autoplot(air) +
  autolayer(fc, series = "Holt's Method", PI = FALSE) +
  autolayer(fc2, series = "Damped Holt's Method", PI = FALSE) +
  xlab("Year") + ylab("Air passengers in Australia (millions)") +
  ggtitle("Forecasts from Holt's method") +
  guides(colour = guide_legend(title = "Forecasts"))
```

The red line shows the forecast using dampning effect and blue one shows forecast with no dampning effect.

Now we will compare the forecast accuracy 3 methods, Simple Exponential Smoothing, Holt's Method and Damped Holt's method.
```{r}
autoplot(livestock) +
  xlab("year") + ggtitle("Livestock, sheep in Asia (millions)")
```

```{r}
e1 = tsCV(livestock, ses, h = 1)
e2 = tsCV(livestock, holt, h = 1)
e3 = tsCV(livestock, holt, damped = TRUE, h = 1)

print ("Compare MSE")
print(c('CV error from SES :-', mean(e1^2, na.rm = TRUE)))
print(c('CV error from Holt :-', mean(e2^2, na.rm = TRUE)))
print(c('CV error from Damped Holt :-', mean(e3^2, na.rm = TRUE)))

print ("Compare MAE")
print(c('CV error from SES :-', mean(abs(e1), na.rm = TRUE)))
print(c('CV error from Holt :-', mean(abs(e2), na.rm = TRUE)))
print(c('CV error from Damped Holt :-', mean(abs(e3), na.rm = TRUE)))

```

Damped Holt's method seems to gives best result on MSE and MAE.

```{r}
fc = holt(livestock, damped = TRUE)
fc[["model"]]
```

Slope parameter is close to zero that means slope of trend is not changing with time.
level parameter is close to one, showing that level reacts strongly to each new observations.

```{r}
autoplot(fc) +
  xlab("Year") + ggtitle("Livestock, sheep in Asia (millions)")
```

The wide prediction interval reflecting the variation in the historical data but the forecast looks reasonable.

