---
title: "5(b). Trend and Seasonal Methods"
author: "Ankit"
date: "2/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Holt's linear trend method :-
This method extends the simple exponential smoothing to allow the forecasting of data with a trend.

This method has 2 smoothing equations and 1 forecast equations:-

**Forecast Equation :- ** $\hat y_{t+h|t} = l_t + hb_t$.

**Level Equation :- ** $l_t = αy_t + (1-α)(l_{t-1} + b_{t-1})$.

**Trend Equation :- ** $b_t = β^* (l_t-t_{t-1}) + (1-β^*)b_{t-1}$.

$l_t$ = estimate of level at time t.
$b_t$ = estimate of trend at time t. 

α is smooting parameter for level and $β^*$ is the smoothing parameter for trend. Both takes values between 0 and 1.

As with simple exponential smoothing, the level $l_t$ is the weighted average of the observation $y_t$ and the one-step training forecast of for time t, here given by $l_{t-1} + b_{t-1}$.
The trend component is the weighted average of estimated trend at time t based on $l_t - l_{t-1}$ and previous estimate of trend $b_{t-1}$.

Example :- Air Passanger data
```{r}
library(fpp2)
autoplot(window(ausair, start = 1990)) +
  xlab("Year") + ylab("") 
```

The timeplot shows that there is an increasing trend in the data.

```{r}
autoplot
```

