---
title: "7(a). Dynamic Regression - Intro"
author: "Ankit"
date: "4/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

ARIMA Models or state space models allows for including information from past observations of a series, but do not allow inclusion of other informations like effect of hoildays, promo activities, competitior activity etc. 

On the other hand, Time series models can take care of these other relevant information but do not include past observations as input.

**Dynamic Regression** is an approach to extend ARIMA models in order to allow other information to be included in the models.
We can write a dynamic regression model as 
$$y_t = \beta_0 + \beta_1 x_{1,t} +.....+ \beta_kx_{k,t} + \eta_t$$
where $\eta_t$ follows an ARIMA model instead of being a white noise.

The challange with these type of models is that we cannot least square kind of methods where we minimise error term (in this case $\eta_t$) to find the estimates. This happens because of several reasons.

1. The estimated coefficients are no longer the best estimates as some information has been ignored, which we have considered in the ARIMA model.

2. As the $\eta_t$ do not follow white noise distribution but an ARIMA Model, all the t-test will be insignificant.

3. AICc or BIC cannot be used in most of the cases to choose the best model.

4. p-values may not be valid in most of the cases. Usually the p-value gets smaller.

Alternatively, we will use maximum likelihood estimation.

An important consideration when estimating a regression with ARMA errors is that all of the variables in the model must first be stationary. Thus, we first have to check that $y_t$ and all of the predictors $(x_{1,t},…, x_{k,t})$ appear to be stationary. If we estimate the model when any of these are non-stationary, the estimated coefficients will not be consistent estimates.  One exception to this is the case where non-stationary variables are co-integrated. If there exists a linear combination of the non-stationary $y_t$ and the predictors that is stationary, then the estimated coefficients will be consistent.

So, we first difference the non-stationary variables. Now, we often need to maintain the form of relationship in between $y_t$ and the predictors, and consequently it is common to difference all the variables if any of them needs differencing. The resulting model is called a "model in differences". 

If all of the variables in the model are stationary, then we only need to consider ARMA errors for the residuals. It is easy to see that a regression model with ARIMA errors is equivalent to a regression model in differences with ARMA errors.

The model above with differencing can be written as
$$y_{t}^{'} = \beta_0 + \beta_1 x_{1,t}^{'} +.....+ \beta_kx_{k,t}^{'} + \eta_{t}^{'}$$
where $y_{t}^{'} =  y_{t} - y_{t-1}$, $x_{t,i}^{'} =  x_{t,i} - x_{t-1,i}$ and $\eta_{t}^{'} = \eta_t - \eta_{t-1}$.

**Example**
US Personal Consumption and Income
```{r}
library(fpp2)
autoplot(uschange[,1:2], facets = TRUE) +
  xlab("Year") + ylab("") +
  ggtitle("Quarterly changes in US consumption
    and personal income")
```

The data looks stationary so there is no need for differencing. We can use auto.arima to fit a dynamic regression model.
```{r}
fit = auto.arima(uschange[,"Consumption"],
                 xreg = uschange[,"Income"])
summary(fit)
```
The fitted model is 
$$y_t = 0.599 + 0.2028x_t + \eta_t$$
$$\eta_t = 0.6922\eta_{t-1} + e_t - 0.5758 e_{t-1} + 0.198 e_{t-2}$$

Now we will plot both ARIMA and regression errors
```{r}
cbind("Regression Errors" = residuals(fit, type = "regression"),
      "ARIMA Errors" = residuals(fit, type = "innovation")) %>% 
  autoplot(facets = TRUE)
```

It is the ARIMA errors that should resemble a white noise series.
```{r}
checkresiduals(fit)
```

