---
title: "6(d). ARIMA vs ETS Models"
author: "Ankit"
date: "3/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The difference between ETS and ARIMA models that all the ETS models are non-stationary but some of the ARIMA models are stationary. 

The ETS models with seasonality or non-damped trend or both have two unit roots (i.e., they need two levels of differencing to make them stationary). All other ETS models have one unit root (they need one level of differencing to make them stationary).

While selecting the models between ARIMA and ETS, we cannot compare the models directly using information criteria because both are calculating the likelihood differenctly.

One way to compare ARIMA and ETS model is time series cross-validation. 

**Non-Seasonal Data**
**Example :- Australian Air traffic data**
```{r}
library(fpp2)
fets = function(x,h){
  forecast(ets(x), h = h)
}

farima = function(x,h){
  forecast(auto.arima(x), h = h)
}

air = window(ausair, start = 1990)

#To compute the CV error
e1 = tsCV(air, fets , h = 1)
e2 = tsCV(air, farima , h = 1)
mean(e1^2, na.rm = TRUE)
mean(e2^2, na.rm = TRUE)
```

ETS model gives smaller error in this particular case. 

```{r}
air %>% ets() %>% forecast() %>% autoplot()
```

**Seasonal Data**
**Quarterly cement production data**
```{r}
cement = window(qcement, start = 1988)
#Preparing trainng set for the model
train = window(cement, end = c(2007,4))

## Fitting ARIMA model on training data
fit.arima = auto.arima(train)
summary(fit.arima)
```

```{r}
checkresiduals(fit.arima)
```

The ARIMA model seems to work well as the residuals looks like white noise.

```{r}
fit.ets = ets(train)
summary(fit.ets)
```

```{r}
checkresiduals(fit.ets)
```

This model also does well in capturing all the dynamics in the data, as the residuals similarly appear to be white noise.

Comparing the 2 models
```{r}
a1 = fit.arima %>% forecast(h = 4*(2013 - 2007) + 1) %>% 
  accuracy(qcement)
a1[,c("RMSE","MAE","MAPE","MASE")]
```

```{r}
a2 = fit.ets %>% forecast(h = 4*(2013 - 2007) + 1) %>% 
  accuracy(qcement)
a2[,c("RMSE","MAE","MAPE","MASE")]
```

Notice that the ARIMA model fits the training data slightly better than the ETS model, but that the ETS model provides more accurate forecasts on the test set.

Forecast from ETS Model 
```{r}
cement %>% ets() %>% forecast(h=12) %>% autoplot()
```


